---
title: "Assignment 5 - K-means"
author: "Samantha Wan"  
format:
  html:
    code-fold: true 
    embed-resources: true
    toc: true
    number-sections: true
    theme: cerulean
---

# Goal of this assignment

The goal of this assignment is for you to **explore the k-means algorithm** learned in class.

In specific, you will be asked to:\
- Use a different number of k than what we did in class,\
- Validate the results of the new model, and\
- Compare your results with those found in class

# Instructions

You will first need to watch (and code along with) the following k-means video:

-   k-means lab: https://youtu.be/GFgMp5tYiMU?si=CI0E-2r-cYZLGVL1 (start from beginning of video, watch till the 01:10)\
-   The partial script for the video above is in our GitHub repository, 04-classcode, **02-24_multivar_kmeans_partial.qmd**. Move the script mentioned above into the `code` subfolder of your `06_multivar` project

**AFTER watching the video and developing the code in class**, then start working on this HW assignment (instructions below): - Move this script into the `code` subfolder of your `06_multivar` project\
- On this script, fill your name under the `title` section of the YAML\
- Go back to the class k-means script, choose a different number of k than what we did in class, and run it.

**Setup**

```{r}
#| message: false
#| warning: false

#install.packages("ggcorrplot")
#install.packages("factoextra")
#install.packages("ggpmisc")

library(tidyverse)
library(ggcorrplot)
library(broom)
library(car)
library(factoextra)
library(ggpmisc)
```

```{r}
weather <- read.csv("../data/weather_monthsum.csv")

head(weather)
```

```{r}
#Selecting only numeric variables and normalizing the data 
weather_n <- weather %>%
  dplyr::select(-c(year:strength_gtex))

weather_norm  <- weather_n %>%
  mutate(across(everything(), ~scale(.x)))
```

# Questions

## 1. What number of k did you select? Why?

```{r}
#| message: false
#| warning: false

# Elbow and Silhouete Methods for visualizing dataset to determine K-value
fviz_nbclust(weather_norm,
             method = "wss",
             k.max = 10,
             FUNcluster = kmeans
             )

fviz_nbclust(weather_norm,
             method = "s",
             k.max = 10,
             FUNcluster = kmeans
             )

```

***I selected 3 k means clusters because based on the two visualization's 3 would be the next best choice to 4 clusters. Based on the Within Sum of Squares (elbow method) we see a more moderate drop between 3 and 4 than we did between 1 & 2 and 2 & 3. Then when looking at the Silhouette method 3 had the second highest peak.***

```{r}
mod_km3 <- kmeans(weather_norm,
                  centers = 3,
                  nstart =10
                  )

mod_km3
```

## 2. How many observations are there in each cluster?

```{r}
weather %>%
  mutate(cluster = mod_km3$cluster) %>%
  group_by(cluster) %>%
  tally() 
```

***There were 108 observations in cluster 1, 253 in cluster 2, and 337 in cluster 3.***

## 3. Using the `fviz_cluster()` function, import here the original plot with k=4 did in class, and the new plot with the number of k you selected for this exercise. How do they visually compare? Which one seems to be a better choice, and why?

```{r}
mod_km4 <- kmeans(weather_norm,
                  centers = 4,
                  nstart =10
                  )
fviz_cluster(mod_km4,
             data = weather_norm
             )
```

```{r}
fviz_cluster(mod_km3,
             data = weather_norm
             )
```

***I still think that 4 was the optimal choice when looking at these clusters using the K means method of clustering. When we look at the visualization of 3 clusters its apparent that there is another cluster to the far left that should be it's own distinct cluster which is what we see when we cluster into 4 groups.***

# Submitting your work

Once you have developed all the code and answers, make sure to Render this quarto file.

**Notes on rendering**:

-   Make sure to render your work and inspect how the final html look like.\
-   If it does not look professional for whatever reason, then fix the issue, re-render it, recheck.\
-   Only send me your work once your html file looks professional.
    -   **DO NOT** delete the file's heading levels (# and ##). They set up the proper heading 1 and 2 levels, and I use them to guide my grading.\

    -   If a given chunk is also outputting warnings or messages, inhibit this behavior by changing the chunk options `message` and `warning` to `FALSE` (I'll demo this in a moment).

    -   If, after rendered, 2 lines of text are connected and you wish to "break line" between them, add 2 extra spaces after the first one (I'lld demo this in a moment).

After rendering, an .html file will be created on your `code` folder.

Rename this file to `Assignment-05-kmeans-LASTNAME.html`.

For ex., mine would be `Assignment-05-kmeans-Bastos.html`.

Submit your work by uploading the **html** file to **eLC Assignment #5 - K-means** by Monday March 2nd 11:59 pm.
